---
title: "LUCAS exclusion rules"
author: 
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
```

In this short note we define some exclusion rules to be applied to the LUCAS sampled and observed points.

IMPORTANT: these exclusion rules can be applied with two different aims:

1. To completely eliminate these observations and then apply the "standard" calibration procedure

2. To use these observations only in a first phase estimation to produce only 1-digit LC and LU estimates, and then apply a second phase to estimate the 2-digit and 3-digit estimates including the 1-digit estimates in the known totals.



We read the clean data, to which have been associated the sampling information (WGT_LUCAS and STRATUM_LUCAS):

```{r}
s1 <- fread("D:/Google Drive/LUCAS 2025/2.DATA/LUCAS22_corrected_complete.csv")
nrow(s1)
```



First, we identify the observations with incomplete Land Cover values by inserting a "x" in the spaces:

```{r}
s2 <- s1
s2$SURVEY_LC1 <- ifelse(nchar(as.character(s2$SURVEY_LC1)) == 2,
                           paste0(s2$SURVEY_LC1,"X"),s2$SURVEY_LC1)
s2$SURVEY_LC1 <- ifelse(nchar(as.character(s2$SURVEY_LC1)) == 1,
                           paste0(s2$SURVEY_LC1,"XX"),s2$SURVEY_LC1)
table(s2$SURVEY_LC1)
```
Then, we flag for elimination all the observations with a "X" (but not those with "B"):

```{r}
s2$elim <- grepl("X",s2$SURVEY_LC1)
s2$elim <- ifelse(substr(s2$SURVEY_LC1,1,1)=="B",FALSE,s2$elim)
table(s2$elim)
```
We have 5,594 cases to be eliminated.

We proceed in the elimination:

```{r}
s2 <- s2[s2$elim == FALSE,]
nrow(s2)
```


Then, we eliminate some values that are not present in the previous rounds of the survey (this elimination should be confirmed):

```{r}
s3 <- s2[!s2$SURVEY_LC1 %in% c("A00",
                               "A10","A20",
                               "B00","B10","B20","B30","B40","B50","B60","B70","B80",
                               "C00",
                               "C20",
                               "D00",
                               "E00",
                               "F00",
                               "G00",
                               "G10","G20",
                               "H00","H10","H20","H30"),]
s3 <- s3[!s3$SURVEY_LU1 %in% c("8","U100","U110",
                               "U200","U220",
                               "U300","U310","U320","U360",
                               "U400"),]
nrow(s3)
```
Only 5 observations have been identified to be eliminated.

Final LC and LU distributions:

```{r}
table(s3$SURVEY_LC1)
```
```{r}
table(s3$SURVEY_LU1)
```


Finally, we analyze the observations to be eliminated:

```{r}
s4 <- s1[!s1$POINT_ID %in% s3$POINT_ID,]
table(s4$SURVEY_LC1)
```

```{r}
table(s4$SURVEY_LU1)
```
We observe the following:

- there is a small incidence of the observations that would be eliminated (1.4%);

- in terms of LC, a prevalence of C2 and C3 values (CONIFEROUS WOODLAND and MIXED WOODLAND)

- in terms of LU, a prevalence of U112 (Fallow land)

We underline that U112 is widely used in the determination of the FAO class, to assign the value "3".

On the basis of the above, it should be decided one of the following choices:

1. to proceed with the standard estimation procedure making use of all the observations, even if some anomalous estimates are produced

2. to proceed with the standard estimation procedure, but without the observations with incomplete values

3. to proceed with the "light" two-phase estimation procedure


